---
title: "synthdid paper results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{synthdid paper results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--html_preserve-->
<style>
pre, pre code, .sourceCode {
  white-space: pre;
  overflow-x: scroll;
  width: 800px;
  max-width: 2400px;
}
</style>
<!--/html_preserve-->

```{r knitr-opts, echo=FALSE}
options(width=999)
```

## Summary
In this vignette, we generate the figures and tables from [Arkhangelsky et al.](https://arxiv.org/abs/1812.09970).
The first section, on the California Proposition 99 application, generates Table 1 and Figure 1.
The second section runs the placebo simulations discussed in Section 3 and then summarizes them by generating Tables 2-4 and Figure 2. 


## The estimators considered
* sdid: synthetic diff-in-diff
* sc  : synthetic control
* did:  diff-in-diff
* mc:   the matrix completion estimator of [Athey et al.](https://arxiv.org/abs/1710.10251)
* difp: de-meaned synthetic control, as proposed in [Doudchenko and Imbens](https://arxiv.org/abs/1610.07748https://arxiv.org/abs/1610.07748) and [Ferman and Pinto](https://arxiv.org/abs/1911.08521).


```{r define-estimators}
library(synthdid)
library(MCPanel)

mc_estimate = function(Y, N0, T0) {
    N1=nrow(Y)-N0
    T1=ncol(Y)-T0
    W <- outer(c(rep(0,N0),rep(1,N1)),c(rep(0,T0),rep(1,T1)))
    mc_pred <- mcnnm_cv(Y, 1-W, num_lam_L = 20)
    mc_fit  <- mc_pred$L + outer(mc_pred$u, mc_pred$v, '+')
    mc_est <- sum(W*(Y-mc_fit))/sum(W)
    mc_est
}
difp_estimate = function(Y, N0, T0) { 
    synthdid_estimate(Y, N0, T0, weights=list(lambda=rep(1/T0, T0)))  
} 

estimators = list(sdid=synthdid_estimate, 
                  sc=sc_estimate, 
                  did=did_estimate, 
                  mc=mc_estimate, 
                  difp=difp_estimate)
```

# California Proposition 99 Application

```{r california-estimation}
set.seed(12345)
data('california_prop99')
setup = panel.matrices(california_prop99)

estimates = lapply(estimators, function(estimator) { estimator(setup$Y, setup$N0, setup$T0) } )
standard.errors = mapply(function(estimate, name) { 
  if(name == 'mc') { NA } else { sqrt(vcov(estimate, method='placebo')) }
}, estimates, names(estimators))
```

## Table 1
```{r california-table}
california.table = rbind(unlist(estimates), unlist(standard.errors))
rownames(california.table) = c('estimate', 'standard error')
colnames(california.table) = toupper(names(estimators))

round(california.table[, c(3,2,1,4,5)], digits=1)
```

## Figure 1. Columns are DID, SC, SDID in that order.

```{r california-trends-plot, fig.width=15, fig.height=5, fig.path="figures/", dev=c("png", "pdf")}
synthdid_plot(rev(estimates[1:3]), facet.vertical=FALSE, 
              control.name='control', treated.name='california', 
              lambda.comparable=TRUE, se.method = 'none', 
              trajectory.linetype = 1, line.width=.75, effect.curvature=-.4,
              trajectory.alpha=.7, effect.alpha=.7, 
              diagram.alpha=1, onset.alpha=.7) + 
    theme(legend.position=c(.26,.07), legend.direction='horizontal', 
          legend.key=element_blank(), legend.background=element_blank(),
          strip.background=element_blank(), strip.text.x = element_blank())
```

```{r california-units-plot, fig.width=15, fig.height=5, fig.path="figures/", dev=c("png", "pdf")}
synthdid_units_plot(rev(estimates[1:3]), se.method='none') + 
    theme(legend.background=element_blank(), legend.title = element_blank(), 
          legend.direction='horizontal', legend.position=c(.17,.07), 
          strip.background=element_blank(), strip.text.x = element_blank())
```

# Placebo Simulations

## load data 

```{r load-penn-and-cps-data}
last.col = function(X) { X[, ncol(X)] }

data(CPS)
Y.logwage      = panel.matrices(CPS, treatment='min_wage', outcome='log_wage', treated.last=FALSE)$Y 
Y.hours        = panel.matrices(CPS, treatment='min_wage', outcome='hours',    treated.last=FALSE)$Y
Y.urate        = panel.matrices(CPS, treatment='min_wage', outcome='urate',    treated.last=FALSE)$Y
w.minwage      = last.col(panel.matrices(CPS, treatment='min_wage',   treated.last=FALSE)$W)
w.gunlaw       = last.col(panel.matrices(CPS, treatment='open_carry', treated.last=FALSE)$W)
w.abortion     = last.col(panel.matrices(CPS, treatment='abort_ban',  treated.last=FALSE)$W)

data(PENN)
Y.loggdp       = panel.matrices(PENN, treatment='dem', outcome='log_gdp', treated.last=FALSE)$Y
w.democracy    = last.col(panel.matrices(PENN, treatment='dem',  treated.last=FALSE)$W)
w.education    = last.col(panel.matrices(PENN, treatment='educ', treated.last=FALSE)$W)

default=list(rank = 4, N1 = 10, T1 = 10)
cps.baseline.params  = estimate_dgp(Y.logwage, w.minwage, default$rank)
```

## Define a function for creating simulators.
A simulator is a no-arg functions returning a simulated dataset, and this function
is used to define them by modifying an extant dgp specification.

* To specify the dgp parameters, we pass in a set of baseline parameters as estimated by estimate.dgp
  and a set of functions `F`, `M`, etc. with names corresponding to those parameters which compute
  the actual parameters for the simulator as functions of the corresponding baseline parameter
* If a numeric value of resample is passed, in each simulation resample that number of units, 
  i.e., rows from `L = M + F`, to use in the simulation design. This is used in some rows of Table 4.

This returns a list with two elements.

1. In the slot `$run`, the simulator, a no-args function returning a simulated dataset.
2. In the slot `$params`, the parameters of the dgp.

```{r define-simulations-setup}
simulator = function(params = cps.baseline.params,
                     F=function(x){x}, M=function(x){x}, 
                     Sigma = function(x){x}, pi = function(x){x}, 
                     ar_coef = function(x){x},
                     N1=default$N1, T1=default$T1, resample=NULL) {
    
    updated.params = list(F=F(params$F), M=M(params$M), 
                          Sigma=Sigma(params$Sigma), pi = pi(params$pi), 
                          ar_coef=ar_coef(params$ar_coef))
    
    list(params=updated.params, N1=N1, T1=T1,
         run=function() { 
             if(!is.numeric(resample)) {
                simulate_dgp(updated.params, N1, T1) 
             } else {       
                ind = sample.int(nrow(updated.params$F), size=resample, replace=TRUE)
                resampled.params=updated.params
                resampled.params$F=updated.params$F[ind,]
                resampled.params$M=updated.params$M[ind,]
                simulate_dgp(resampled.params, N1, T1) 
            } 
        })
}    
```

## Define simulators.
* Simulators use `Sigma`, not `ar_coef`, to generate noise, so updating `ar_coef` doesn't affect the simulations. 
* We do update `ar_coef` below so `$params` has the correct coefficients. These will be shown in Table 2.

```{r define-simulations}
simulators = list(
    baseline   =  simulator(),
    # Modified outcome model
    no.corr    =  simulator(Sigma=function(Sigma) { diag(nrow(Sigma)) * norm(Sigma,'f')/sqrt(nrow(Sigma))}, 
                            ar_coef=function(coefs) { 0*coefs }),
    no.M       =  simulator(M=function(M) { 0*M }), 
    no.F       =  simulator(F=function(F) { 0*F }),
    only.noise =  simulator(M=function(M) { 0*M }, F=function(F) {0*F}),
    no.noise   =  simulator(Sigma=function(Sigma) { 0*Sigma }, ar_coef=function(coefs) { 0*coefs }),
    # Modified assignment process
    gun.law    =  simulator(estimate_dgp(Y.logwage, w.gunlaw,   default$rank)), 
    abortion   =  simulator(estimate_dgp(Y.logwage, w.abortion, default$rank)),
    random     =  simulator(pi=function(pi) { rep(.5,length(pi)) }),
    # Modified outcome variable
    hours      =  simulator(estimate_dgp(Y.hours,   w.minwage,  default$rank)),
    urate      =  simulator(estimate_dgp(Y.urate,   w.minwage,  default$rank)),
    # Assignment block size
    T1.is.one  = simulator(T1=1),
    N1.is.one  = simulator(N1=1),
    blocksize.is.one = simulator(T1=1, N1=1),
    # Resample rows (from Table 4)
    resample.200 = simulator(resample=200, N1=20),
    resample.400 = simulator(resample=400, N1=40),
    # penn world table (Table 3)
    penn.democracy  = simulator(estimate_dgp(Y.loggdp,  w.democracy, default$rank)),
    penn.education  = simulator(estimate_dgp(Y.loggdp,  w.education, default$rank)),
    penn.random     = simulator(estimate_dgp(Y.loggdp,  w.education, default$rank), 
                                pi=function(pi) { rep(.5,length(pi)) }))
```

## Run simulations.
* Do many point estimates, as these are fast, to estimate rmse and bias for Tables 2 and 3.
* Do fewer standard error estimates, as these are slow, for coverage for Table 4.
* Save results in replications/replication*.RDS. Assumes you've created the directory ./replications

### Details
This is a long-running computation, taking roughly 3 days on a recent macbook air. It is written to save data as it runs, one file per replication, so it can be continued if interrupted. To continue, just rerun the loop: only replications for which these files are missing are run. 

The loop over replications gives each replication its own RNG seed, computed as 
a function of the initial seed 12345. Within each replication, every call to
a simulator, estimator, or variance estimator uses this replication-specific seed:
we explicitly restore the seed before each call. As a result, changing the set of 
simulations, estimators, or variance estimators considered does not change the results we get
for any specific simulator/estimator/variance-estimator within a given replication. Nor does 
interrupting and continuing the loop. 

By using nextRNGStream to choose replication-specific seeds for the L'Ecuyer-CMRG RNG, 
we get streams of random numbers that are more or less independent from replication to replication.
See `help(nextRNGStream)`.

```{r set-up-simulations}
run.replication = function(seed, coverage.for=c()) { 
    library(synthdid)
    library(MCPanel)
    
    estimate = array(NA, dim = c(length(simulators), length(estimators), 4))
    for (ss in 1:length(simulators)) {
        set.seed(seed)
        setup = simulators[[ss]]$run()
        for (ee in 1:length(estimators)) {
            set.seed(seed)
            est = estimators[[ee]](setup$Y, setup$N0, setup$T0)
            estimate[ss, ee, 1] = est
            if (names(estimators)[ee] %in% coverage.for) {
                methods = c('bootstrap', 'jackknife', 'placebo')
                for (mm in 1:length(methods)) {
                  set.seed(seed)
                  estimate[ss, ee, mm + 1] = sqrt(vcov(est, methods = methods[mm]))
                }
            }
        }
    }
    
    estimate
}
```


```{r run-simulations}
library(parallel)
library(doParallel)

sim.replications  = 1000
coverage.replications = 400
coverage.estimators = c('sdid', 'sc', 'did')
cores = 8

# set up RNG seeds for each replication
RNGkind("L'Ecuyer-CMRG")
set.seed(12345)
seeds = as.list(rep(NA, sim.replications))
seeds[[1]] = .Random.seed
for(rr in 2:length(seeds)) { seeds[[rr]] = nextRNGStream(seeds[[rr-1]]) }

# set up workers
cl = makeCluster(cores, outfile='simulations.log')
registerDoParallel(cl, cores=cores)

# run simulations
estimates.list = foreach(rr = 1:sim.replications) %dopar% { 
    RNGkind("L'Ecuyer-CMRG")
    filename = sprintf('replications/replication%d.RDS', rr)
    if(!file.exists(filename)) {
        start.time = Sys.time()
        estimate = run.replication(seeds[[rr]], coverage.for =
            if(rr <= coverage.replications) { coverage.estimators } else { c() })
        saveRDS(estimate, file=filename)
        end.time = Sys.time()
        print(sprintf('replication %d ran for %1.0f hours/minutes/seconds, from %s to %s', 
                      rr, end.time-start.time, start.time, end.time))
        estimate
    }
}

# shut down workers
stopCluster(cl)
```

## Load estimates from disk and concatenate them into an array
`estimates` will be a #simulations x #estimators x #estimands x #replications array where the 4 estimands are 

1. the point estimate
2. the bootstrap estimate of standard error
3. the jackknife estimate of standard error
4. the placebo estimate of standard errror

```{r load-estimates}
library(abind)

estimates.list = foreach(rr = 1:sim.replications) %do% {
    readRDS(sprintf('replications/replication%d.RDS', rr))
}
estimates = do.call(abind, c(estimates.list, list(rev.along=0))) 
```

## Compute summary statistics from simulations

```{r summary-stats}
rmse = apply(estimates, c(1,2), function(x) { sqrt(mean(x[1,]^2)) })
bias = apply(estimates, c(1,2), function(x) { mean(x[1,]) })
coverage = lapply(2:4, function(se) {
    apply(estimates[,,,1:coverage.replications],
          c(1,2), function(x) { mean(abs(x[1,] / x[se,]) <= qnorm(.975))  })
})
```

## Point Estimation: Tables 2 and 3
### Compute summary info about simulator designs to include in Table 2.

```{r simulator-info}
sim.info = do.call(rbind, lapply(simulators, function(sim) { 
    data.frame(F.scale  =  sqrt(mean(sim$params$F^2)), 
               M.scale  =  sqrt(mean(sim$params$M^2)),
               noise.scale  =  sqrt(mean(diag(sim$params$Sigma))),
               ar.lag1      =  sim$params$ar_coef[1],
               ar.lag2      =  sim$params$ar_coef[2])
}))
```

## Display a point estimate summary table 
In the paper, this is split into Tables 2 and 3, with CPS sims in the former and PENN sims in the latter. Also save in point_table.tab.

```{r point-estimate-table}
point.table = cbind(sim.info, as.data.frame(cbind(rmse, bias)))
rownames(point.table) = names(simulators)
colnames(point.table) = c('F', 'M', 'Sigma', 'ar:lag1', 'ar:lag2', 
                          sprintf('rmse:%s', names(estimators)), 
                          sprintf('bias:%s', names(estimators)))
write.table(round(point.table,2),  file='point_table.tab')
round(point.table, digits=3)
```

## Display a coverage summary table. 
A subset of these rows are in Table 4 of the paper. Also save in coverage_table.tab.

```{r coverage-table}
coverage.table = do.call(cbind, coverage)
rownames(coverage.table) = names(simulators)
colnames(coverage.table) = c(sprintf('bootstrap:%s',  names(estimators)),
                             sprintf('jackknife:%s',  names(estimators)),
                             sprintf('placebo:%s',    names(estimators)))
keep = !is.na(coverage.table[1,])

write.table(round(coverage.table[,keep],2), file='coverage_table.tab')
round(coverage.table[,keep], digits=2)
```

## Plot the error distribution of the estimates (Figure 2)
We plot an estimate of error density function for the baseline CPS setting and the minimum wage assignment variant. This is Figure 2. 


* Recall estimates has dimensions (simulator, estimator, estimand, replication) with estimand=1 corresponding to the point estimate. 
* The estimate=2,3,4 slots are corresponding standard error estimates.

```{r error-density-setup}
grid = expand.grid(ss=1:length(simulators), ee=1:length(estimators))
error.densities = do.call(rbind, mapply(function(ss,ee) {
    errors = estimates[ss,ee,1,]
    density.estimate = lindsey_density_estimate(errors, K = 100, deg = 3) 
    data.frame(x=density.estimate$centers, 
               y=density.estimate$density,
               simulator=names(simulators)[ss],
               estimator=names(estimators)[ee])
    }, grid$ss, grid$ee, SIMPLIFY=FALSE))
```

```{r error-density-plot, fig.width=8, fig.height=4, fig.path="figures/", dev=c("png", "pdf")}
library(ggplot2)

show.sims       = error.densities$simulator %in% c('baseline', 'random')
show.estimators = error.densities$estimator %in% c('did', 'sc', 'sdid')
# reformat the information in $simulator and $estimator for display in the plot title and legend
error.densities$Title = ifelse(error.densities$simulator == 'baseline', 
                               'Minimum Wage Assignment', 'Random Assignment')
error.densities$Estimator = toupper(error.densities$estimator)

ggplot(error.densities[show.sims & show.estimators, ]) + 
    geom_line(aes(x=x,y=y,color=Estimator)) + geom_vline(xintercept=0, linetype=3) + 
    facet_wrap(~Title) + xlab('estimate') + ylab('density') + 
    theme_light() + theme(legend.position=c(.94,.88), 
                          legend.background=element_blank(), 
                          legend.title=element_blank()) 
```
