---
title: "plotting features"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plotting features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(synthdid)
library(ggplot2)
```

# Dates as Timesteps
Sometimes we want to use dates as timesteps.
We store timesteps as the colnames of Y, and colnames can't be Date objects,
so to work around this we can use strings convertible to dates via as.Date.
Functions in the package, like the plotting functions, are aware of this convention, 
so the dates get displayed correctly, with correct spacing is retained when intervals vary.
```{r, fig.width=7.5, fig.height=4}
  data(california_prop99)
  california_prop99$date = as.Date(sprintf('%04d/%02d/%02d', california_prop99$Year, 1, 1))
  setup = panel.matrices(california_prop99[! california_prop99$year %in%  c(1974:1977, 1985:1993),], time='date')
  estimate = synthdid_estimate(setup$Y, setup$N0, setup$T0)
  plot(estimate)
```

# Spaghetti Plots
We can include the trajectories of individual controls on our parallel trends diagram.
Here we plot the 10 controls with the largest weights $\omega_i$. 
``{r, fig.width=7.5, fig.height=4}
  top.controls = synthdid_controls(estimate)[1:10, , drop=FALSE]
  plot(estimate, spaghetti.units=rownames(top.controls))
``

# Complex Plots
We can even overlay multiple estimates in one plot and include a magnified view of some regions.
We create a flipbook in which we bring synthetic diff-in-diff's control
closer to California's trajectory to we can see how parallel it is clearly, overlay it completely
so we can better compare it to the ADH synthetic control method's synthetic control,
plot the ADH synthetic control estimate, and zoom in on a region around treatment onset
in which we can see the ADH synthetic control possibly begin to diverge from California's trajectory.

Because we display multiple estimates, we cannot use the overlay parameter to slide synthetic
diff-in-diff's control toward California. Instead, we pass this parameter as an attribute
'overlay' of the estimator, which overrides the function argument and is interpreted the same way.
Including an invisible (via alpha.multiplier) copy of tau.hat keeps the axes of the plot
from changing as the visible one's control trajectory moves.
---
# This example requires the right mix of packages and libraries to knit properly
# into an animation. To just plot 5 pictures, replace the next line with:
# {r, fig.width=7, fig.height=4.5}
---
```{r, fig.width=7, fig.height=4.5, fig.show='animate', ffmpeg.format='gif', dev='jpeg'}
estimate.sc = sc_estimate(setup$Y, setup$N0, setup$T0)
with.overlay = function(est, s) { attr(est,'overlay') = s; est } 
estimators = function(s) { 
  estimator.list = list(with.overlay(estimate, s), estimate.sc, estimate) 
  names(estimator.list)=c('Synthetic Diff-in-Diff', 'Synthetic Control', '')
  estimator.list
}

plot.estimators = function(ests, alpha.multiplier) {
  synthdid_plot(ests, se.method='none', 
                alpha.multiplier=alpha.multiplier, facet=rep(1,length(ests)),
	        trajectory.linetype = 1, effect.curvature=-.4,
	        trajectory.alpha=.5, effect.alpha=.5, diagram.alpha=1) + 
    scale_alpha(range=c(0,1), guide='none') 
}

# set up the box we zoom in on in plot 5
time = as.integer(timesteps(setup$Y))
xbox.ind = c(which(lambda(estimate) > .01)[1], setup$T0+4)
xbox = time[xbox.ind] + c(-.5,.5)
ybox = range(setup$Y[setup$N0+1, min(xbox.ind):(max(xbox.ind))]) + c(-4,4)


p1 = plot.estimators(estimators(0),   alpha.multiplier=c(1,.1,0))  
p2 = plot.estimators(estimators(.75), alpha.multiplier=c(1,.1,0)) 
p3 = plot.estimators(estimators(1),   alpha.multiplier=c(1,.1,0)) 
p4 = plot.estimators(estimators(1),   alpha.multiplier=c(1, 1,0)) 
p4.zoom = p4 + coord_cartesian(xlim=xbox, ylim=ybox) + xlab('') + ylab('') + 
    theme(axis.ticks.x= element_blank(), axis.text.x = element_blank(), 
          axis.ticks.y=element_blank(), axis.text.y=element_blank(), 
          legend.position='off')
p5 = p4 + annotation_custom(ggplotGrob(p4.zoom), xmin = 1968,   # location by
                            xmax = 1984.7,   ymin=2, ymax=95) + # trial and error
	        geom_rect(aes(xmin=min(xbox), xmax=max(xbox), 
	                      ymin=min(ybox), ymax=max(ybox)), 
	                  color=alpha('black', .25), size=.3, fill=NA)


plot.theme = theme(legend.position=c(.9,.85), legend.direction='vertical', 
                   legend.key=element_blank(), legend.background=element_blank())

p1 + plot.theme
p2 + plot.theme
p3 + plot.theme
p4 + plot.theme
p5 + plot.theme
```
