---
title: "synthdid introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{synthdid introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
ragg_png = function(..., res = 192) {
  ragg::agg_png(..., res = res, units = "in")
}
knitr::opts_chunk$set(dev = "ragg_png", fig.ext = "png")
```

```{r setup}
library(synthdid)
library(ggplot2)
set.seed(12345)
```

# Example use 
We estimate the effect of California Proposition 99 on cigarette consumption

## Standard Errors
We have only one treated unit (California), so the only implemented method to estimate
the standard error is the 'placebo' method described in Section 5 of Arkhangelsky et al. 
Because this is not trustworthy, it is not used by default: vcov instead returns NA.
Here it is probably too large. 

But to show how confidence intervals are constructed and displayed, we'll use it anyway.
To do this, we pass method='placebo' to vcov and pass se.method='placebo' 
to our plotting functions so they do the same. 
If we didn't, the plots wouldn't show have standard errors.
```{r}
data('california_prop99')
setup = panel.matrices(california_prop99)
tau.hat = synthdid_estimate(setup$Y, setup$N0, setup$T0)

tau.hat
print(summary(tau.hat))

se = sqrt(vcov(tau.hat, method='placebo'))
sprintf('point estimate: %1.2f', tau.hat)
sprintf('95%% CI (%1.2f, %1.2f)', tau.hat - 1.96 * se, tau.hat + 1.96 * se)
```

## The parallel trends plot. 
  We show the trajectories of California  and the synthetic control we compare it to, 
  with a vertical line indicating onset of treatment. The relative size of the 
  weights $\lambda_t$ for pre-treatment periods are shown on the bottom. 
  Overlaid is a diagram of the estimate. 
  
### The Parallelogram
  The blue segment shows the change from the weighted pre-treatment average 
  to the post-treatment average California. The red one does the same for
  a synthetic control. 
  
### The Hallucination
  Our premise is that, absent treatment, California would change like
  the synthetic control: from California's pre-treatment average,
  we draw a dashed segment parallel to the control's red one.
  Where it ends is our estimate of what we'd have observed in California had, 
  contrary to fact, it not been treated. 
  
### The Treatment Effect
  Comparing this counterfactual estimate to what we observed in 
  the real world, we get a treatment effect. We show it with a black arrow. 
  We show a 95% confidence interval with two gray arrows: 
  one for the upper confidence bound and one for the lower.

```{r fig.width=7, fig.height=4}
plot(tau.hat, se.method='placebo')
```

## The control unit contribution plot. 
  We could draw the plot above, and get a questionable treatment effect estimate,
  by comparison to any individual control state. Our actual estimate is a weighted average
  of these: the weights are the synthetic control weights $\omega_i$. This plot
  shows each individual comparison with a dot, with weight $\omega_i$ determining the dot size.
  The black horizontal line shows the actual effect; 
  the gray ones show the endpoints of a 95% confidence interval.

```{r fig.width=7, fig.height=4}
synthdid_units_plot(tau.hat, se.method='placebo')
```

## Checking for pre-treatment parallel trends
When the synthetic control and treated trajectories are far from one another,
it can be hard to see how parallel they are. Pass overlay=1 to plot to overlay them. 

```{r fig.width=7, fig.height=4}
plot(tau.hat, overlay=1,  se.method='placebo')
```
Pass values between zero and one to shift the control's trajectory
toward California.
```{r fig.width=7, fig.height=4}
plot(tau.hat, overlay=.8, se.method='placebo')
```

# Compare to other estimators
We compare to synthetic control and diff-in-diff. 
```{r}
tau.sc   = sc_estimate(setup$Y, setup$N0, setup$T0)
tau.did  = did_estimate(setup$Y, setup$N0, setup$T0)
estimates = list(tau.did, tau.sc, tau.hat)
names(estimates) = c('Diff-in-Diff', 'Synthetic Control', 'Synthetic Diff-in-Diff')
```

```{r}
print(unlist(estimates))
```

```{r fig.width=7, fig.height=4}
synthdid_plot(estimates, se.method='placebo')
```
```{r fig.width=15, fig.height=5}
synthdid_units_plot(estimates, se.method='placebo')
```

# Customize plots 
Arguments to plotting functions can change the transparency, line thickness, etc. of individual elements.
The output of every plotting function is a ggplot, so you can change theme, legend position, etc. as usual.
We'll use these to match the style of Figure 1 in Arkhangelsky et al.
```{r fig.width=15, fig.height=5}
synthdid_plot(estimates, facet.vertical=FALSE, 
              control.name='control', treated.name='california', 
              lambda.comparable=TRUE, se.method = 'none', 
              trajectory.linetype = 1, line.width=.75, effect.curvature=-.4,
              trajectory.alpha=.7, effect.alpha=.7, 
              diagram.alpha=1, onset.alpha=.7) + 
    theme(legend.position=c(.26,.07), legend.direction='horizontal', 
          legend.key=element_blank(), legend.background=element_blank(),
          strip.background=element_blank(), strip.text.x = element_blank())

synthdid_units_plot(estimates, se.method='none') + 
    theme(legend.background=element_blank(), legend.title = element_blank(), 
          legend.direction='horizontal', legend.position=c(.17,.07), 
          strip.background=element_blank(), strip.text.x = element_blank())
```
